---
title: "Basic Treatment Response Model"
output: html_notebook
---

___
Required Imports
```{r}
library("Rcpp")
library("tidyverse")
library("devtools")
library("brms")
library("rstan")
library("rstantools")
library("tidybayes")
```


___
make a data frame that has two columns the first with “treatment” with values 
1 through 5 and the second with “response” with random gaussian values with a different mean for each treatment. So there are say 100 rows for each treatment, so the data frame frame has 500 rows total.

```{r}
data <- tibble::tibble(treatment = as.factor(1:5)) %>% # piping tibble(treatment) into mutate. tibble constructs a df
  dplyr::mutate( # allows you to manipulate dataframe
    treatment_mean = rnorm( # finding the treatment mean
      n = 5,
      mean = 10,
      sd = 1)) %>% # piping treatment mean into rowwise
  dplyr::rowwise() %>% # piping rowwise into do 
  dplyr::do({ # do applies a function to each group (in this case rowwise)
    treatment_params <- . # the df
    tibble::tibble( # constructing a df
      treatment = treatment_params$treatment, # putting treatment into the df
      treatment_mean = treatment_params$treatment_mean, # treatment_mean into the df
      value = rnorm( # response value
        n = 100, # using 100 random Gaussian distribution values around the mean
        mean = treatment_params$treatment_mean,
        sd = 3)) # with a sd as 3
  }) %>%
  dplyr::ungroup() # removes grouping done by group_by 
```

___
Summary of the values in data

```{r}
data %>%
  dplyr::group_by(treatment) %>%
  dplyr::summarize(
    treatment_mean = treatment_mean[1],
    empirical_mean = mean(value),
    sd = sd(value))
```

___
Plot of treatment and response

```{r}
ggplot(data = data,
       aes(x = treatment,
           y = value)) +
  geom_point(size = 1.2,
             alpha = 0.8) + 
  theme_minimal() +
  labs(title = "Response of Each Treatment",
       x = "Treatment",
       y = "Response")
```
___
Fitting the data to a simple Bayesian model (brm)
formula: 
1. A dependent variable we want to predict
2. ~ that is used to indicate that we now give the other variables of interest.
3. insert that the dependent variable has a variance and that we want an intercept.

value (dependent variable) ~ 0 (intercept) + treatment (variance)

```{r}
model <- brms::brm(
  value ~ 0 + treatment,
  data = data)
```
___
Model info
```{r}
model
```
___
Running model_w_prior

```{r}
model_w_prior <- brms::brm(
  formula = value ~ 0 + treatment, 
  data = data,
  prior = c(
    brms::prior(normal(10,3), 
                class = "b",
                lb = 0), 
    brms::prior(normal(3,0.5),
                class = "sigma"
                ) )
)
```
___
Model_w_prior info
```{r}
model_w_prior

```
___
Graphing prior and posterior distribution
```{r}
prior_posterior_plot <- function(model,irq = .95) {
  
  model_prior <- model %>%
    brms:::update.brmsfit(sample_prior = "only") # updates the sample priors in model

  draws <- dplyr::bind_rows( # bind rows returns a df
    model %>% # piping model into tidydraws
      tidybayes::tidy_draws() %>% # extract draws from a Bayesian fit into a wide format data frame
      tidybayes::gather_variables() %>% # gathers variables from a tidy dataframe of draws from variables into a single column
      dplyr::mutate(sample_type = "Posterior"), # mutating sample type to read posterior
    model_prior %>%
      tidybayes::tidy_draws() %>% 
      tidybayes::gather_variables() %>%
      dplyr::mutate(sample_type = "Prior")) %>% # renaming sample type 
    dplyr::filter(!stringr::str_detect(.variable, "__$")) %>% # detecting a string pattern in a variable
    dplyr::group_by(.variable) %>% # grouping by variable
      dplyr::filter( # filtering values based on quantile
        .value < quantile(.value, 1 - (1-irq)/2),
        .value > quantile(.value, (1 - irq)/2)) %>%
    dplyr::ungroup() # removing grouping done by group_by 

  ggplot2::ggplot(data = draws) + # graphing draws
    ggplot2::theme_bw() + # black and white theme
    ggplot2::theme(legend.position = "bottom") + # legend position
    ggplot2::geom_density( # smooth density estimates
      mapping = ggplot2::aes( # aesthetic mapping 
        x = .value, # what is being plotted
        # y = log(1 + ..density..),
        fill = sample_type, # fill done by sample type
        group = sample_type), # grouped by sample type
      color = "black", # outline color is black 
      alpha = .9) +   # transparency  
    ggplot2::ggtitle( # plot title
      label = paste0("Prior vs Posterior Marginal Distribution: ", model$name, " model")) +
    ggplot2::facet_wrap( # makes a 1d sequence of panels into 2d
      facets = dplyr::vars(.variable), # the number of panels = # variables
      scales = "free") + 
    ggplot2::scale_y_continuous("Density") + # y label along the whole y axis
    ggplot2::scale_x_continuous("Parameter Value") + # x label along the whole x axis
    ggplot2::scale_fill_discrete("Distribution") # scale is discrete for all plots
}
```
```{r}
prior_posterior_plot(model_w_prior)
```
___
Less data
```{r}
data2 <- tibble::tibble(treatment = as.factor(1:5)) %>% # piping tibble(treatment) into mutate. tibble constructs a df
  dplyr::mutate( # allows you to manipulate dataframe
    treatment_mean = rnorm( # finding the treatment mean
      n = 5,
      mean = 10,
      sd = 1)) %>% # piping treatment mean into rowwise
  dplyr::rowwise() %>% # piping rowwise into do 
  dplyr::do({ # do applies a function to each group (in this case rowwise)
    treatment_params <- . # the df
    tibble::tibble( # constructing a df
      treatment = treatment_params$treatment, # putting treatment into the df
      treatment_mean = treatment_params$treatment_mean, # treatment_mean into the df
      value = rnorm( # response value
        n = 50, # using 50 random Gaussian distribution values around the mean
        mean = treatment_params$treatment_mean,
        sd = 3)) # with a sd as 3
  }) %>%
  dplyr::ungroup() # removes grouping done by group_by 
```
___
model2 fitting
```{r}
model2 <- brms::brm(
  value ~ 0 + treatment,
  data = data)
```
___
model2 info
```{r}
model2
```
less data, greater error
___

smaller sd
```{r}
data3 <- tibble::tibble(treatment = as.factor(1:5)) %>% # piping tibble(treatment) into mutate. tibble constructs a df
  dplyr::mutate( # allows you to manipulate dataframe
    treatment_mean = rnorm( # finding the treatment mean
      n = 5,
      mean = 10,
      sd = 1)) %>% # piping treatment mean into rowwise
  dplyr::rowwise() %>% # piping rowwise into do 
  dplyr::do({ # do applies a function to each group (in this case rowwise)
    treatment_params <- . # the df
    tibble::tibble( # constructing a df
      treatment = treatment_params$treatment, # putting treatment into the df
      treatment_mean = treatment_params$treatment_mean, # treatment_mean into the df
      value = rnorm( # response value
        n = 100, # using 100 random Gaussian distribution values around the mean
        mean = treatment_params$treatment_mean,
        sd = runif(n = 100, min = 0, max = 1))) # with a sd between 0 & 1 (default)
  }) %>%
  dplyr::ungroup() # removes grouping done by group_by 
```
```{r}

model3 <- brms::brm(
  value ~ 0 + treatment,
  data = data)

```
___
model3 info
```{r}
model3
```
smaller sd, slightly smaller error
